Global patterns of forest autotrophic carbon fluxes

### 2020-08-05 manuscript/manuscript draft.Rmd
slightly condensed results

In total, we analyzed 1319 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years. These records represented a total of **#** plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Fig. 1, Table 2).

@@

All major carbon fluxes decreased linearly with latitude (Fig. 2; Table S2). Latitude was a strong predictor for many of the carbon fluxes, particularly the larger fluxes (Table S2). Specifically, latitude explained 64% of variation in GPP (n = 243, p<0.0001), 50% in NPP (n = 161, p<0.0001) and 44% in ANPP (n = 278, p<0.0001). The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18). The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, which were best explained by the logarithmic model.

@@

All fluxes increased with MAT (all p<0.05; Figs. 3-4, S4-S5, Table S2). 
For eight of the nine fluxes, this relationship was linear.
For only one variable, $BNPP$, did a lognormal fit provide signficant improvement over a first-order linear relationship. 
*As with latitude, MAT tended to explain more variation in the larger flux ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this -- NB these values are correct**

MAP was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 25% of variation in C flux. 
For five of the nine fluxes, a second-order polynomial model provided the best fit, while first-order linear and lognormal relationships each provided the best fit for two flxues. 
All fluxes increased with MAP up to at least 2000 mm, above which responses were variable (Figs. 4, S4-S5).

There was a significant additive effect of MAT and MAP on $GPP$, $ANPP$ and $R_{auto}$ (Fig. 3, Table S3), and a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{stem}$ (Fig. 3, Table S3). The interaction was negative for $NPP$ and positive for $ANPP_{stem}$. For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, MAP did not have a signficant effect when accounting for MAT (Fig. 3, Table S3).

@@

Our results indicated that annual forest C fluxes were most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of C fluxes. In addition to MAT, several of its correlates (Fig. S2) were consistently identified as strong univariate predictors of C fluxes: temperature seasonality, annual temperature range, annual frost days, PET, and length of growing season (Figs. 4, S4-S7).

All C flux variables were significantly related to potential evapotranspiration. $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ showed a logarithmic relationship with PET; however, all other fluxes showed a polynomial relationship with PET (Fig. 4c, S4-5; Table S2). We found strong evidence for a saturation point or peak with PET: C fluxes tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was also evidence that some C fluxes begin to decrease at values above 1800mm PET.

@@

We found a significant relationship between length of growing season and C fluxes, with all fluxes showing a linear increase with length of growing season (Figs. 4e, S6-S7; Table S2). Length of growing season was a strong predictor of C fluxes, explaining 53% of variation in GPP, 38% of variation in NPP, and 34% of variation in ANPP (all p<0.05; Talbe S2), but it was a weaker predictor than MAT for all fluxes analysed (Table S4).

@@

When annual C fluxes were standardized by growing season length (in monthly increments), correlations with growing season climate were generally weak (Figs. S8-S9). Specifically, $ANPP$ increased with growing season temperature ($R^2$ = 0.09, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). Growing season solar radiation had a positive influence on $BNPP$ ($R^2$ = 0.17, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.13, p<0.01). Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.07, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.10, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05). All other relationships were non-significant.

@@

Our analysis of a large global database (ForC) reveals how autotrophic carbon fluxes in mature forests vary with latitude and climate on a global scale. We show that, across all nine forest autotrophic C fluxes analyzed, C cycling decreases continually with latitude (*H1.1*; Fig. 2)--a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals that of tropical forests [@huston_global_2009]. C fluxes generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig. S2) and with component fluxes summing appropriately to larger fluxes (Fig. 2), indicating no major, systematic omissions or overestimations of flux components. However, climate explained lower proportions of variability among subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP_{fine.root}$, $R_{root}$; Fig. 2; Table S2).

Latitudinal variation in C fluxes is primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs. 3-4). Water availability is also influential, but generally of secondary importance across the climate space occupied by forests (Figs. 3-4). Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of C fluxes (*H4*; Figs. 4e-f, S2, S6-S7), though growing season length did not improve upon MAT as a predictor. Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of C fluxes (*H5*; Fig. S9; Table S4).

>>>>

In total, we analyzed 1319 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years. These records represented a total of 255 plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Fig. 1, Table 2).

@@

All major carbon fluxes decreased with latitude (Fig. 2; Table S2). Latitude was a strong predictor for many of the carbon fluxes, particularly the larger fluxes (Table S2). Specifically, latitude explained 64% of variation in GPP (n = 243, p<0.0001), 50% in NPP (n = 161, p<0.0001) and 44% in ANPP (n = 278, p<0.0001). The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18). The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, which were best explained by the logarithmic model.

@@

All fluxes increased with MAT (all p<0.05; Figs. 3-4, S4-S5, Table S2). 
For eight of the nine fluxes, this relationship was linear.
For only one variable, $BNPP$, did a lognormal fit provide signficant improvement over a first-order linear relationship. 
*As with latitude, MAT tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this -- NB these values are correct**

MAP was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 25% of variation in C flux. All fluxes increased with MAP up to at least 2000 mm, above which responses were variable (Figs. 4, S4-S5).   
There was a significant additive effect of MAT and MAP on $GPP$, $ANPP$ and $R_{auto}$ (Fig. 3, Table S3), and a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{stem}$ (Fig. 3, Table S3). The interaction was negative for $NPP$ and positive for $ANPP_{stem}$. For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, MAP did not have a signficant effect when accounting for MAT (Fig. 3, Table S3).

@@

Our results indicated that annual forest C fluxes were most strongly explained by temperature-related climate variables at the global scale. In addition to MAT, several of its correlates (Fig. S2) were consistently identified as strong univariate predictors of C fluxes: temperature seasonality, annual temperature range, annual frost days, PET, and length of growing season (Figs. 4, S4-S7).

All C flux variables showed a significant relationship with potential evapotranspiration. The relationship was logarithmic for $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$, and polynomial for all other fluxes (Fig. 4c, S4-5; Table S2). We found strong evidence for a saturation point or peak with PET: C fluxes tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was also evidence that some C fluxes begin to decrease at values above 1800mm PET.

@@

We found a significant relationship between length of growing season and C fluxes, with all fluxes showing a positive relationship with length of growing season (Figs. 4e, S6-S7; Table S2). Length of growing season was a strong predictor of C fluxes, explaining 53% of variation in GPP, 38% of variation in NPP, and 34% of variation in ANPP (all p<0.05; Table S2), but it was a weaker predictor than MAT for all fluxes analysed (Table S4).

@@

When annual C fluxes were standardized by growing season length (in monthly increments), correlations with growing season climate were generally weak (Figs. S8-S9). $ANPP$ increased with growing season temperature ($R^2$ = 0.09, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). Growing season solar radiation had a positive influence on $BNPP$ ($R^2$ = 0.17, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.13, p<0.01). Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.07, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.10, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05). All other relationships were non-significant.

@@

Our analysis of a large global database (ForC) reveals how autotrophic carbon fluxes in mature forests vary with latitude and climate on a global scale. We show that, across all nine forest autotrophic C fluxes analyzed, C cycling decreases continually with latitude (*H1.1*; Fig. 2)--a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals that of tropical forests [@huston_global_2009]. C fluxes generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig. S2) and with component fluxes summing appropriately to larger fluxes (Fig. 2), indicating no major, systematic omissions or overestimations of flux components. However, climate explained lower proportions of variability among subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP_{fine.root}$, $R_{root}$; Fig. 2; Table S2). 
Latitudinal variation in C fluxes is primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs. 3-4). Water availability is also influential, but generally of secondary importance across the climate space occupied by forests (Figs. 3-4). Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of C fluxes (*H4*; Figs. 4e-f, S2, S6-S7), though growing season length did not improve upon MAT as a predictor. Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of C fluxes (*H5*; Fig. S9; Table S4).
